The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-04-05 src/ms.tex
Improve consistency of formatting software names

Sphinx is a common and useful tool for generating documentation for Python
packages. The Astropy project maintains a default Sphinx configuration along
with Astropy-specific extensions that can be easily added to community
projects via the
\href{https://github.com/astropy/sphinx-astropy}{\texttt{sphinx-astropy}}
meta-package. This tool provides a pre-configured Sphinx setup compatible with
Astropy projects, which includes several extensions useful for generating API
documentation
(\href{https://github.com/astropy/sphinx-automodapi}{\texttt{sphinx-automodapi}}),
allowing for \package{numpy} docstring parsing
(\href{https://github.com/numpy/numpydoc}{\texttt{numpydoc}}), embedded image
handling
(\href{https://github.com/sphinx-gallery/sphinx-gallery}{\texttt{sphinx-gallery}};
\href{https://github.com/python-pillow}{\texttt{pillow}}), advanced
documentation testing support
(\href{https://github.com/astropy/pytest-doctestplus}{\texttt{pytest-doctestplus}}),
and providing a custom documentation theme ideal for analysis packages
(\href{https://github.com/astropy/sphinx-astropy}{\texttt{astropy-sphinx-theme}}).

@@

Software developers at the Space Telescope Science Institute (STScI),
the operations center of JWST, have been developing \python-based tools
for JWST since 2010 (starting with the JWST Calibration Reference Data
System) and have provided major contributions to \astropy from its
inception. The JWST instrument calibration pipelines, exposure-time
calculators, and data analysis tools are all written in Python and
depend on the \astropypkg core package and some coordinated and
affiliated packages. For the \astropypkg core package, the JWST mission
has provided extensive contributions to the \astropysubpkg{modeling},
\astropysubpkg{units}, \astropysubpkg{coordinates}, \astropysubpkg{wcs},
\astropysubpkg{io.fits}, \astropysubpkg{io.votable},
\astropysubpkg{stats}, \astropysubpkg{visualization}, and
\astropysubpkg{convolution} subpackages as well as to the general
package infrastructure and maintenance.

@@

A prototype for the low-level analysis is \texttt{ctapipe} \citep{ctapipe},
a \python package developed to perform all the necessary tasks to perform data processing, from the raw data
of Cherenkov telescopes to the reconstructed event lists.
The high-level analysis (or CTA science tools) will be based on the Astropy affiliated package
Gammapy~\citep{gammapy}.
Both Gammapy and \texttt{ctapipe} make heavy use of the \astropy core package, mainly for units, times, coordinate transformations,
tables, and FITS I/O.
As CTA will record gamma-ray events with a rate of up to $10\,000$ events per second,
it needs to perform a large number of coordinate transformations.
To enable this, CTA member M.~Nöthe contributed a major performance improvement
for large numbers of coordinates with different observation times,
based on earlier work by B.~Winkel.
Together with Gammapy maintainer A.~Donath and former maintainer C.~Deil,
a total of 107 merged pull requests were contributed to astropy.

@@

The focus of Learn Astropy over the coming months and years is to further
improve the interface for sharing educational content on the main Learn Astropy
website, and to facilitate the development of new content that highlights
\astropy and \astropy-affiliated package functionality.
However, we simultaneously also plan to solicit indexing and ingestion of
third-party tutorial series, to provide a unified interface for identifying
astronomy-specific educational content that demonstrates the functionality
available in the vast ecosystem of open-source software packages.
We also plan to look for opportunities to expand the reach of \astropy workshops
beyond the AAS meetings.

@@

Furthermore, the \astropy packages would not exist in their current form without
a number of web services for code hosting, continuous integration, and
documentation; in particular, \astropy heavily relies on GitHub, Azure
Pipelines, CircleCI, and Read the Docs.

@@

\astropypkg interfaces with the SIMBAD database, operated at CDS, Strasbourg,
France. It also makes use of the ERFA library \citep{erfa}, which in turn
derives from the IAU SOFA Collection\footnote{\url{http://www.iausofa.org}}
developed by the International Astronomical Union Standards of Fundamental
Astronomy \citep{sofa}.

>>>>

Sphinx is a common and useful tool for generating documentation for \python
packages. The Astropy project maintains a default Sphinx configuration along
with Astropy-specific extensions that can be easily added to community
projects via the
\href{https://github.com/astropy/sphinx-astropy}{\texttt{sphinx-astropy}}
meta-package. This tool provides a pre-configured Sphinx setup compatible with
Astropy projects, which includes several extensions useful for generating API
documentation
(\href{https://github.com/astropy/sphinx-automodapi}{\texttt{sphinx-automodapi}}),
allowing for \package{numpy} docstring parsing
(\href{https://github.com/numpy/numpydoc}{\texttt{numpydoc}}), embedded image
handling
(\href{https://github.com/sphinx-gallery/sphinx-gallery}{\texttt{sphinx-gallery}};
\href{https://github.com/python-pillow}{\texttt{pillow}}), advanced
documentation testing support
(\href{https://github.com/astropy/pytest-doctestplus}{\texttt{pytest-doctestplus}}),
and providing a custom documentation theme ideal for analysis packages
(\href{https://github.com/astropy/sphinx-astropy}{\texttt{astropy-sphinx-theme}}).

@@

Software developers at the Space Telescope Science Institute (STScI),
the operations center of JWST, have been developing \python-based tools
for JWST since 2010 (starting with the JWST Calibration Reference Data
System) and have provided major contributions to \astropy from its
inception. The JWST instrument calibration pipelines, exposure-time
calculators, and data analysis tools are all written in \python and
depend on the \astropypkg core package and some coordinated and
affiliated packages. For the \astropypkg core package, the JWST mission
has provided extensive contributions to the \astropysubpkg{modeling},
\astropysubpkg{units}, \astropysubpkg{coordinates}, \astropysubpkg{wcs},
\astropysubpkg{io.fits}, \astropysubpkg{io.votable},
\astropysubpkg{stats}, \astropysubpkg{visualization}, and
\astropysubpkg{convolution} subpackages as well as to the general
package infrastructure and maintenance.

@@

A prototype for the low-level analysis is \texttt{ctapipe} \citep{ctapipe},
a \python package developed to perform all the necessary tasks to perform data processing, from the raw data
of Cherenkov telescopes to the reconstructed event lists.
The high-level analysis (or CTA science tools) will be based on the Astropy affiliated package
\package{gammapy}~\citep{gammapy}.
Both \package{gammapy} and \texttt{ctapipe} make heavy use of the \astropy core package, mainly for units, times, coordinate transformations,
tables, and FITS I/O.
As CTA will record gamma-ray events with a rate of up to $10\,000$ events per second,
it needs to perform a large number of coordinate transformations.
To enable this, CTA member M.~Nöthe contributed a major performance improvement
for large numbers of coordinates with different observation times,
based on earlier work by B.~Winkel.
Together with \package{gammapy} maintainer A.~Donath and former maintainer C.~Deil,
a total of 107 merged pull requests were contributed to astropy.

@@

The focus of Learn Astropy over the coming months and years is to further
improve the interface for sharing educational content on the main Learn Astropy
website, and to facilitate the development of new content that highlights
\astropypkg and \astropy-affiliated package functionality.
However, we simultaneously also plan to solicit indexing and ingestion of
third-party tutorial series, to provide a unified interface for identifying
astronomy-specific educational content that demonstrates the functionality
available in the vast ecosystem of open-source software packages.
We also plan to look for opportunities to expand the reach of \astropy workshops
beyond the AAS meetings.

@@

Furthermore, the \astropy packages would not exist in their current form without
a number of web services for code hosting, continuous integration, and
documentation; in particular, \astropy heavily relies on \github, Azure
Pipelines, CircleCI, and Read the Docs.

@@

\astropypkg interfaces with the SIMBAD database, operated at CDS, Strasbourg,
France. It also makes use of the \package{ERFA} library \citep{erfa}, which in turn
derives from the IAU SOFA Collection\footnote{\url{http://www.iausofa.org}}
developed by the International Astronomical Union Standards of Fundamental
Astronomy \citep{sofa}.
